Weakly Nonlinear MRI: Notes

### 2014-03-31 notes/mri_notes.tex
second order
complete second order
NEW amplitude equation

%% Graphics and math
\usepackage{graphics,graphicx,subfigure}
\usepackage{amsmath}

@@

$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_x\Pi \, + \, \frac{1}{4\pi} B_0\partial_z^3A \, + \, \frac{1}{4\pi}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{1}{4\pi}\partial_zA\partial_x\partial_z^2 A \, - \, \partial{1}{4\pi}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{1}{4\pi}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\

@@

$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\

@@

$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0\partial_z\partial_x^2 A \, + \, J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\

@@

$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\

@@

$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\

@@

$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - J(A, \nabla^2 A)\\
J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
-J(A, \Psi) \\
J(\Psi, B_{1y}) - J(A, u_{1y}) \\
\end{matrix} \right]$ \\

@@

$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\

@@

$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, J(A_1, \nabla^2 A_1) $ \\

@@

$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \widetilde{J}(A_1, \nabla^2 A_1) $ \\

@@

>>>>

%% Graphics and math
\usepackage{graphics,graphicx,subfigure}
\usepackage{amsmath}
\usepackage{amsfonts}

@@

$\partial_t \partial_z^2 \Psi \, + \, \partial_z^2\Psi\partial_x\partial_z\Psi \, + \, \partial_z\Psi \partial_x\partial_z^2 \Psi \, - \, \partial_z\partial_x\Psi\partial_z^2\Psi \, - \, \partial_x \Psi \partial_z^3 \Psi \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_z\partial_x\Pi \, + \, \frac{1}{4\pi} B_0\partial_z^3A \, + \, \frac{1}{4\pi}\partial_z^2 A \partial_x \partial_z A \, + \, \frac{1}{4\pi}\partial_zA\partial_x\partial_z^2 A \, - \, \frac{1}{4\pi}\partial_z\partial_x A \partial_z^2 A \, - \, \frac{1}{4\pi}\partial_xA\partial_z^3A \, + \, \frac{1}{\reye} \nabla^2 \partial_z^2 \Psi $ \\

@@

$\partial_t\partial_z^2\Psi \, + \, J\left(\Psi, \partial_z^2\Psi\right) \, - \, 2\partial_z u_{1y} \, = \, \frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0 \partial_z^3A \, + \, \frac{1}{4\pi}J\left(A, \partial_z^2A\right) \, + \, \frac{1}{\reye}\nabla^2 \partial_z^2 \Psi$ \\

@@

$\partial_t \partial_x^2 \Psi \, + \, J\left(\Psi, \partial_x^2 \Psi\right) \, = \, -\frac{1}{4\pi}\partial_x\partial_z\Pi \, + \, \frac{1}{4\pi}B_0\partial_z\partial_x^2 A \, + \, \frac{1}{4\pi}J\left(A, \partial_x^2 A \right) \, + \, \frac{1}{\reye} \nabla^2 \partial_x^2 \Psi$ \\

@@

$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, 2 \partial_z u_{1y} \, = \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\

@@

$\partial_t \nabla^2 \Psi \, + \, J\left(\Psi, \nabla^2 \Psi\right) \, - \, \frac{1}{4\pi}J\left(A, \nabla^2 A \right) \, =  \, 2 \partial_z u_{1y} \,+ \, \frac{1}{4\pi} B_0 \partial_z \nabla^2 A \, + \, \frac{1}{\reye}\nabla^4 \Psi$ \\

@@

$\mathbf{N} = \left[ \begin{matrix}  J(\Psi, \nabla^2 \Psi) - \frac{1}{4\pi}J(A, \nabla^2 A)\\
J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y}) \\ 
-J(A, \Psi) \\
J(\Psi, B_{1y}) - J(A, u_{1y}) \\
\end{matrix} \right]$ \\

@@

$N^{(\Psi)} \to \epsilon^2 J(\Psi_1, \nabla^2 \Psi_1) \,-\, \epsilon^2 \frac{1}{4\pi}J(A_1, \nabla^2 A_1) \,+\, \epsilon^3 J(\Psi_1, \nabla^2\Psi_2) \, - \epsilon^3\frac{1}{4\pi}J(A_1, \nabla^2 A_2) \,+\, \epsilon^3 J(\Psi_2, \nabla^2\Psi_1) \, - \, \epsilon^3 \frac{1}{4\pi}J(A_2, \nabla^2 A_1) \,+\, 2\epsilon^3 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2\epsilon^3 \frac{1}{4\pi}J(A_1, \partial_z\partial_Z A_1) \,+\, \epsilon^3\widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \, \epsilon^3 \frac{1}{4\pi}\widetilde{J}(A_1, \nabla^2 A_1) \, + \, \mathcal{O}(\epsilon^4)$ \\

@@

$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, \frac{1}{4\pi}J(A_1, \nabla^2 A_1) $ \\

@@

$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, \frac{1}{4\pi}J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, \frac{1}{4\pi}J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 \frac{1}{4\pi}J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \frac{1}{4\pi}\widetilde{J}(A_1, \nabla^2 A_1) $ \\

@@

The solution to the $\mathcal{O}(\epsilon)$ equation is: \\

$\mathbf{V_1} = \alpha(T, Z) \mathbb{V}_{11} e^{i Q z} + c.c. + \beta(T, Z)\mathbb{U}_{11}$ \\

Where $\alpha$ and $\beta$ are amplitudes, $\mathbb{V}_{11}(x) = (\Psi_{11}, u_{11}, A_{11}, B_{11})^\mathbf{T}$ and $\mathbb{U}_{11} = (0, 0, 0, 1)^\mathbf{T}$. \\

To find the $\mathcal{O}(\epsilon^2)$ solution, must write out $\mathbf{N_2}$. \\

We find: \\

$\mathbf{N_2} = \alpha^2\mathbb{N}_{22} e^{i2Qz} + \left|\alpha\right|^2 \mathbb{N}_{20} + c.c.$ \\

Then: \\

$\mathbf{V_2} \, = \, \alpha^2 e^{i2Qz} \mathcal{L}^{-1}\mathbb{N}_{22} \, + \,  \left|\alpha\right|^2 \mathcal{L}^{-1} \mathbb{N}_{20}\, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ + c.c.\\

And $\partial_Z \mathbf{V_1} = \partial_Z \alpha \mathbb{V}_{11} e^{i Q z} \, + \, \partial_Z \beta \mathbb{U}_{11} \, + \, c.c.$, so we end up with: \\

$\mathbf{V_2} \, = \, \alpha^2 e^{i2Qz} \mathbb{V}_{22} \, + \, \left|\alpha\right|^2 \mathbb{V}_{20} \, - \, \partial_Z \alpha \mathbb{V}_{21} e^{iQz} \, - \, \partial_Z \beta \mathbb{U}_{20} + c.c.$ \\

$\mathbb{V}_{22} = \mathcal{L}^{-1}\mathbb{N}_{22}$ \\

$\mathbb{V}_{20} = \mathcal{L}^{-1} \mathbb{N}_{20}$ \\

$\mathbb{V}_{21} = \mathcal{L}^{-1} \widetilde{\mathcal{L}}_1 \mathbb{V}_{11}$ \\

$\mathbb{U}_{20} = \mathcal{L}^{-1} \widetilde{\mathcal{L}}_1 \mathbb{U}_{11}$ \\

Note that this is the same as Umurhan+ (B27) except for two signs. \\

Our $\mathcal{O}(\epsilon^3)$ equation is: \\

Taking only the terms which are resonant with $e^{iQz}$, we find: \\

$\mathcal{D} \mathbb{V}_{11} e^{iQz}\partial_T \alpha \, + \, \alpha \left| \alpha^2 \right| \mathbf{N}_{31} e^{iQz} \, + \, \alpha \partial_Z \beta \widetilde{\mathbf{N}}_{31} e^{iQz} \, = \, - \partial_Z^2 \alpha \widetilde{\mathcal{L}}_1 \mathbb{V}_{21} e^{iQz} \, + \, \partial_Z^2 \alpha \widetilde{\mathcal{L}}_2 \mathbb{V}_{11} e^{iQz} \, - \, \partial_Z \alpha \mathbb{V}_{11}\mathcal{X}e^{iQz} \, - \, iQ^3 \alpha \mathcal{L}_3 \mathbb{V}_{11} e^{iQz}$ \\

Note that $\mathcal{L}\mathbf{V}_3$ does not enter here.\\

Following Umurhan+'s lead and taking the inner product with $\mathbb{V}^\dagger$, I find the amplitude equation: \\

$a \partial_T \alpha \, + \, c \alpha \left|\alpha^2\right| \, + \, \widetilde{c} \alpha \partial_Z \beta \, = \, - b \partial_Z \alpha \, + \, h \partial_Z^2 \alpha \, - \, g i Q^3 \alpha$ \\

$a \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right>$ \\

$c \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right>$ \\

$\widetilde{c} \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right>$ \\

$b \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right>$ \\

$h \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right>$ \\

$g \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right>$ \\
